---
title: "Intro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(FARS.package)
```

This package is used to perform specific analyis on the Fatality Analyis Reporting System. Data inlcuded with the package is throught years 2013-2015. Analysis possible inlcude development of tibble that allows comparison of fatalities by year, and, development of state map with locations of all fatalities plotted for a year.

##Loading FARS data
While the FARS dataset is broad, the analysis in this package use only a few parameters available.

For more info on the data, see the [NHTSA FARS Manuals & Documentation page](https://crashstats.nhtsa.dot.gov/#/DocumentTypeList/4). 

###Make File Name
The make_filename function takes a numeric as input and returns a character string which serves as a filename for that year.

``` {r echo=TRUE}
fars_2013 <- make_filename(2013)
fars_2013
```
This is then used in conjuction with other functions in the package for analysis

###Single Years
For single years, use the fars_read with the output of make_filename as the input, then pipe that to fars_read_years().

``` {r echo=TRUE}
FARS_14 <- make_filename(2014)%>%
  fars_read()%>%
  fars_read_years()
```

###Multiple Years
The pacakge is really centred on the functions fars_summarize_years, and fars_map_state, each of which have the first three functions, make_filename, fars_read, and, fars_read_years, embedded within them.
Hence, in order to assess multiple years and to create maps of all fatal collisions within a state, one need only use these functions as is, without having to perform the tasks described above.

fars_summarize_years() will return a tibble with the month as rows and the years as columns, and the number of fatal auto collisions during those months and years listed.

``` {r multiple years, echo=TRUE}
fars_summarize_years(2013:2015)
```

fars_map_state() will return a plot of the locations of all fatal auto collisions for a state during that year.

```{r state plot, ech0=TRUE}
fars_map_state(12, 2013)
```






